---
html_document:
  pdf_document: default
  html_notebook: default
author: "Brian Zebosi"
date: "9/17/2021"
output:
  html_document:
    df_print: paged
title: "STAT575_HW_1"
---


 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


 

```{r, include= F}
# Install Required Packages
package_mult_bz <- function(packages){
  not_installed_yet <- packages[!(packages %in% rownames(installed.packages()))]
  if (any(not_installed_yet == TRUE)) {
    install.packages(packages[not_installed_yet, dependencies = TRUE])
  }
  sapply(packages, require, character.only = TRUE)
}

packages <- c("reshape2" , "ggpubr" , "grid" , "multcomp" , "car" , "Rmisc", 
              "ggsignif" , "readxl" , "dendsort" , "viridis" , "circlize" , "tidyverse",
              "ComplexHeatmap", "emmeans" , "lme4" , "Matrix" , "lmerTest","matrixStats","Hmisc","GGally" )

package_mult_bz(packages)
```

### Question_1.

```{r}
x_1<-matrix(c(2,-1,6,-2,5,1,6,5,1,9,2,8,5,8,
            1,0,5,0,4,-2,3,3,-3,7,-4,1,0,1,
            0,2,8,-1,9,7,6,6,-3,5,-5,6,3,5),nrow=7,byrow=F)
```

## 1a.

```{r}
# No. rows
n<- nrow(x_1); n
```

```{r}
# No. columns
p<- ncol(x_1); p
```

## 1b.

```{r}
x_1[3,2]
```

## 1c.

```{r}
colMeans(x_1)
```
## 1d.

```{r}
cov(x_1)
```

## 1e.

```{r}
cor(x_1)
```

## 1f.

```{r}
t(x_1)
```

## 1g.

```{r}
scale(x_1)
```

## 1h. 

```{r}
cov(scale(x_1))
```

```{r}
cor(scale(x_1))
```


The standardized Covariance matrix and correlation matrix are equal.


### Question_2.


```{r}
liver.df <- read.csv("liver_enzymes.csv")
```
## 2a.

```{r}
colMeans(liver.df[,2:4])
```
## 2b. 

```{r}
cov(liver.df[,2:4])
```

## 2c. 

```{r}
cor(liver.df[,2:4])
```
In the correction matrix; x2 and x3, have significant postive correlations.


## 2d 

```{r}
pairs(liver.df[,2:4], panel = panel.smooth, pch=19)
```




Not all pairwise relationships are linear. X1 - X3, and X1 - X2 relationships look curvilinear maybe due to outliers.

## 2e. 

```{r}
det(cov(liver.df[,2:4]))
```
The generalized sample variance is large suggesting that there are variables with very large Eigenvalues.
Large generalized sample variance suggests uncorrelated variables. 

Generalized sample variance is a measure of the total variance, which determines the normal distribution. 

## 2f. 

```{r}
sum(diag(cov(liver.df[,2:4])))
```




### Question_4

```{r}
cereal.df <- read.csv("cereal.csv")
```

## 4a. 

```{r}
cereal.df %>% group_by(manufacturer) %>% tally() -> tallies
tallies
```
## 4b. 

```{r}
# before cleaning
dim(cereal.df) 
## after cleaning
cereal.df %>% filter_if(is.numeric, all_vars(. >= 0)) %>% dim()
```
There are 12 cereals with missing information.

Thus has (12/17) = 15.5% proportion. 

## 4c.

```{r}

cereal.df %>% filter_if(is.numeric, all_vars(. >= 0)) -> cereal2.df
divide_cups <- function(x, na.rm=T)(x/cereal2.df$cups)
cereal2.df %>% select(-c(shelf,weight)) %>% 
  mutate_if(is.numeric, divide_cups) -> cereal3.df
cereal3.df %>% mutate(shelf = cereal2.df$shelf) -> cereal3.df
head(cereal3.df)
```
## 4d.


i).

Highest numbers of calories per cup 

```{r}
cereal3.df$name[cereal3.df$calories==max(cereal3.df$calories)]
```

Lowest numbers of calories per cup 

```{r}
cereal3.df$name[cereal3.df$calories==min(cereal3.df$calories)]
```


ii).

Highest numbers of vitamins per cup 

```{r}
cereal3.df$name[cereal3.df$vitamins==max(cereal3.df$vitamins)]
```

Lowest numbers of vitamins per cup 

```{r}
cereal3.df$name[cereal3.df$vitamins==min(cereal3.df$vitamins)]
```

iii).

```{r}
variable <- c("calories", "carbs", "fat","fiber", "sodium","potassium")
cereal3.df %>% select(variable) %>% pairs(panel = panel.smooth)
```



The cereals that are outliers with respect to calories only are

```{r}
cereal3.df$name[cereal3.df$calories >300]
```

The cereals that is outlier with respect to carbs only is:


```{r}
cereal3.df$name[cereal3.df$carbs >50]
```
The cereals that is outliers with respect to both calories and carbs is:


```{r}
cereal3.df$name[cereal3.df$carbs >50 & cereal3.df$calories >300]
```

calories and potassium, carbs  and fiber, fat and potassium, calories and carbs, calories and fat, calories and fiber show a linear correlation


iv)


```{r}
cereal3.df$shelf <- as.factor(cereal3.df$shelf)
ggplot(cereal3.df, aes(y = sugar, x = as.factor(shelf), fill = as.factor(shelf))) + geom_boxplot() + theme_bw()
```




Distribution of sugar on shelf 3 had greater variation.  

Distribution at shelf 1 was right skewed with low median i.e. lowest sugar per cur.

Distribution of sugar on shelf 2 displayed small variation with outliers and also highest sugar per cup.



v)


```{r}

ggplot(cereal3.df, aes(y = fiber, x = as.factor(shelf), fill = as.factor(shelf))) + geom_boxplot() + theme_bw()
```





Distribution of fiber per cup on shelf 3 is more variable and has the highest fibre per cup.  

Distribution at shelf 1 was right skewed

Distribution of sugar on shelf 2 displayed small variation with outliers and also lowest fibre per cup.


vi)


```{r}
ggparcoord(cereal3.df, columns = 4:12, groupColumn ="shelf") + theme_bw()
```



Based on the parallel coordinate plot, shelf 1 have the highest content or healthier cereals than 2 and 3. 


vii)


```{r}
cereal3_KG <- filter(cereal3.df, manufacturer =="K" | manufacturer =="G")

ggparcoord(cereal3_KG, columns = 4:12, groupColumn ="manufacturer") + theme_bw()
```




Generally Kellogg Mills cereals have higher nutrition than General Mills cereals. 

General mills have lower carbs content and higher fat content compared to Kellogg cereals.

Kellogg mills have more fiber, protein, potassium. 